---
title: "Creating social network"
---

# Set up

```{r}
# Load packages
library(tidyverse)
library(readxl)
library(writexl)
library(arrow)
library(splitstackshape)
library(misty)
library(gtools)
library(igraph)
library(qgraph)

# Set to not show numbers in scientific notation
options(scipen = 100)
```

# Read in data

Read in shs_shcs_cleaned data.

```{r}
shs_shcs_cleaned <- readRDS("../../Data/shs_shcs_cleaned.R")
```

Use uncount to create multiple rows for each household, based on weight.
Use single year fuel poverty weight after housing costs, since this excludes households with missing incomes.

```{r}
# The weights have decimal points, so round these
shs_shcs_cleaned <- shs_shcs_cleaned %>%
  mutate(weight_rounded = round(fpWghtAHC, digits = 0))

# Now use this to expand the dataset
shs_shcs_expanded <- shs_shcs_cleaned %>%
  uncount(weight_rounded)
```

Create property size and income decile/percentile variables for stratification and model running

```{r}
shs_shcs_expanded <- shs_shcs_expanded %>%
  mutate(property_size = case_when(total_floor_area_m2 <= 72 ~ "small",
                                   total_floor_area_m2 <= 98 ~ "medium",
                                   TRUE ~ "large"),
         income_decile = ntile(income_ahc, 10),
         income_ahc_percentile = ecdf(income_ahc)(income_ahc))
```

```{r}
set.seed(1)
stratified_sample_half <- stratified(shs_shcs_expanded,
                                group = c("location",
                                           "income_decile",
                                           "property_size",
                                           "is_off_gas_grid",
                                           "property_type",
                                           "built_form",
                                           "heating_system",
                                           "epc_rating",
                                          "potential_epc_rating",
                                           "occupant_type",
                                           "is_solid_wall",
                                          "roof_energy_efficiency",
                                          "walls_energy_efficiency",
                                          "windows_energy_efficiency",
                                           "is_heat_pump_suitable_archetype"),
                                size = 0.005)
```

Create unique ID

```{r}
stratified_sample_half <- stratified_sample_half %>%
  ungroup() %>%
  mutate(id_2 = c(1:nrow(stratified_sample_half)),
         id = paste0(id, "_", id_2))
```

Create social network

```{r}
set.seed(3)
small_world_half <- sample_smallworld(1, 12513, 2, 0.4)
```

```{r}
deg <- table(degree(small_world_half))

rel_frequency <- deg/sum(deg)

barplot(rel_frequency)
```

```{r}
edge_density(small_world_half)
```

```{r}
isolated <- which(degree(small_world_half) == 0)

no_unconnected <- delete_vertices(small_world_half, isolated)
```

```{r}
smallworldness(no_unconnected, B = 10)
```


Add names from dataframe to sample

```{r}
 V(small_world_half)$name <- unique(stratified_sample_half$id)
```

Get lists of neighbours

```{r}
small_world_half_df <- as_data_frame(small_world_half)
```

```{r}
neighbours <- small_world_half_df %>%
  group_by(from) %>%
  summarise(neighbours = str_c(to, collapse = ','))
```

Add neighbours to sample dataframe and drop unneeded variables

```{r}
stratified_sample_half <- stratified_sample_half %>%
  merge(neighbours, by.x = "id", by.y = "from", all = TRUE) %>%
  select(-gpawr2c, -gpawr3cy_08, -hedqual8, -hhtype_new, -nssec2, -hih_rel, -HIH_ETH2012, -hihage)
```

Write to Parquet file

```{r}
write_parquet(stratified_sample_half, "../../Data/scottish_stratified_sample_1_with_network_half.parquet")
```



