---
title: "R Notebook"
output: html_notebook
---

```{r}
ew_distribution <- read_xlsx("../../Data/hou01tables2021.xlsx", sheet = "4a", skip = 3)
```

Rename columns

```{r}
colnames(ew_distribution) <- c("area_code", "area_name", "tenure", "detached_house", "semi_detached_house", "terraced_house", "flat", "caravan")
```



```{r}
ew_pivot <- ew_distribution %>%
  filter(str_detect(area_code, "E0") | str_detect (area_code, "W06")) %>% # Filter out anything that isn't a local authority
  select(-area_code, -caravan) %>%
  pivot_longer(cols = c(detached_house:flat), values_to = "number", names_to = "type") %>%
  pivot_wider(id_cols = c(area_name, type), names_from = tenure, values_from = number) %>%
  select(area_name, type, "Total: Owned or shared ownership", "Total: Social Rented", "Private rented from landlord or letting agency", "Other private rented")
```

```{r}
colnames(ew_pivot) <- c("area_name", "type", "owner_occupied", "rented_social", "private_landlord", "private_other")
```

Missing values are from suppressed small numbers under 10. Set these at 5.

```{r}
ew_pivot <- ew_pivot %>%
  mutate(private_landlord = as.numeric(private_landlord),
         private_other = as.numeric(private_other),
    owner_occupied = as.numeric(owner_occupied),
         rented_social = as.numeric(rented_social)) %>%
  replace(is.na(.), 5)
```


```{r}
ew_pivot <- ew_pivot %>%
  mutate(rented_private = private_landlord + private_other) %>%
  select(-private_landlord, -private_other)
```

```{r}
ew_pivot_percent <- ew_pivot %>%
  pivot_longer(cols = c(owner_occupied:rented_private), names_to = "tenure", values_to = "number") %>%
  mutate(combined = paste(area_name, type, tenure, sep = "_"),
         percent = number/sum(number)) %>%
  select(combined, percent)
```

Create a combined column for the household agents dataset (remember to come back and write code to read this in)

```{r}
household_agents_for_sample <- household_agents %>%
  mutate(house_type = ifelse(str_detect(built_form, "terrace"), "terraced", built_form),
         house_flat = ifelse(property_type == "house" | property_type == "bungalow", paste0(house_type, "_house"), property_type),
      combined = paste(location, house_flat, occupant_type, sep = "_"))
```

Merge with column with sampling weights

```{r}
household_agents_for_sample_merge <- household_agents_for_sample %>%
  merge(ew_pivot_percent, all.x = TRUE)
```

Any NAs in percent?

```{r}
sum(is.na(household_agents_for_sample_merge$percent))
```

No, great!

So, take sample, with percent column for probabilities

```{r}
set.seed(1)
eng_strat_sample_1 <- slice_sample(household_agents_for_sample_merge,
                             n = 24810,
                             weight_by = percent)
```

Save sampling file as parquet. Also save sample as Parquet.

```{r}
write_parquet(household_agents_for_sample_merge, "../../Data/ew_agents_for_sampling.parquet")
```

```{r}
write_parquet(eng_strat_sample_1, "../../Data/ew_strat_sample_1.parquet")
```

Create a loop to create 9 more samples and save them as parquet files

```{r}
sample_no <- c(2:10)

for (i in sample_no) {
  set.seed(sample_no)
ew_strat_sample <- slice_sample(household_agents_for_sample_merge,
                             n = 24810,
                             weight_by = percent)
write_parquet(ew_strat_sample, paste0("/../../Data/ew_stratified_sample_", i, ".parquet"))
}
```
